---
title: "Practical Machine Learning Project week 4"
author: "Soham De"

date: 23th Oct 2019
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig.width=10, fig.height=5)
options(width=120)
library(lattice)
library(ggplot2)
library(plyr)
library(caret)
library(corrplot)
library(randomForest)
```

# Building model


We do the the following steps:

- Preprocessing  the data given(downloaded)
- Exploratory data analysis on the test data set
- Select a model which is right with the predictons
- Predict the classification of the model on the given test set

## Pre-processing


```{r}
# we read the csv from the downloded files, 2 files one is for training other for testing
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
# create a partition with the training dataset 
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)
```

Both created datasets have 160 variables. Those variables have plenty of NA, so that can be removed with the cleaning procedures below. The Near Zero variance (NZV) variables are also removed and the ID variables also as well.

```{r}

NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TrainSet)

```
```{r}
dim(TestSet)
```


```{r}
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TrainSet)
```

```{r}
dim(TestSet)
```
# remove identification from only variables (columns 1 to 5)
```{r}
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
```

# Correlation Analysis performed

```{r}
corMatrix <- cor(TrainSet[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "lower", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))
```
# Random forest model is applied and trained 
```{r}

set.seed(211)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=controlRF)
modFitRandForest$finalModel
```

```{r}
# prediction on Test dataset which we have downloaded
predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest
```


# plot the new matrix results
```{r}
plot(confMatRandForest$table, col = confMatRandForest$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMatRandForest$overall['Accuracy'], 4)))
```

# Prediction of the following models, in this case random forest is good
```{r}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```

